---
title: "Using the fic R package for focused model comparison: skew-normal models"
date: "`r Sys.Date()`"
author: Chris Jackson <chris.jackson@mrc-bsu.cam.ac.uk>
output: 
  html_document: 
    includes:
      before_body: fic.sty
  code_folding: show
vignette: >
 %\VignetteIndexEntry{Using the fic R package for focused model comparison: skew-normal models}
 %\VignetteEngine{knitr::rmarkdown}
 %\VignetteDepends{fic,ggplot2}
 %\VignetteEncoding{UTF-8}
bibliography: fic.bib
---


Illustrates focused model comparison for a novel class of models. 

This needs narrow model pars to be in the middle! 

Mixture of covariate selection and flexibility selection.  Covariate coefficients are subset of model parameters 



```{r}

library(sn)
data(ais)
plot(density(ais$Hc))
plot(ais$BMI, ais$Hc, pch=19)

ldsnorm <- function(x, mean, sd, skew){
    log(skew) + (skew-1)*pnorm(x, mean, sd, log=TRUE) + dnorm(x, mean, sd, log=TRUE)
}

mloglik <- function(b0, b1, sd, skew){
    ret <- -sum(ldsnorm(ais$Hc, b0 + b1*ais$BMI, sd, skew))
    ret
}

snlm <- function(reg=FALSE, skew=FALSE){
    inilm <- lm(Hc ~ BMI, data=ais)
    ini <- c(coef(inilm), log(summary(inilm)$sigma), 0)
    if (!reg && !skew){
        ini <- ini[c(1,3)]
        fn <- function(par) mloglik(par[1], 0, exp(par[2]), 1)
    } else if (reg && !skew){
        ini <- ini[c(1,2,3)]
        fn <- function(par) mloglik(par[1], par[2], exp(par[3]), 1)
    } else if (!reg && skew){
        ini <- ini[c(1,3,4)]
        fn <- function(par) mloglik(par[1], 0, exp(par[2]), exp(par[3]))
    } else if (reg && skew){
        fn <- function(par) mloglik(par[1], par[2], exp(par[3]), exp(par[4]))
    }
    opt <- nlm(fn, ini, hessian=TRUE)
    vcov <- solve(opt$hessian)
    list(loglik=-opt$minimum, est=opt$estimate, vcov=vcov, nobs=nrow(ais))
}

```

Four models 

1. constant mean, constant variance, standard normal error 

2. linear regression, no skewness ie standard normal error 

3. no regression term, skewed error 

4. linear regression term, skewed error 


```{r}


(mod1 <- snlm(reg=FALSE, skew=FALSE))
(mod2 <- snlm(reg=TRUE, skew=FALSE))
(mod3 <- snlm(reg=FALSE, skew=TRUE))
(mod4 <- snlm(reg=TRUE, skew=TRUE))


```


## Mean at particular cov value, shouldn't need skewness to be known. 


```{r}

focus <- function(par, X){
    X %*% par[1:2]
}

fns <- list(coef=function(x)x$est,
            vcov=function(x)x$vcov,
            nobs=function(x)x$nobs)

med.bmi <- rbind(male=c(1, 23.56), female=c(1, 21.82))

inds <- rbind(c(1,0,1), c(1,1,1))
fic(mod2, inds=inds, inds0=c(1,0,1), fns=fns, focus=focus, X=med.bmi,
    sub=list(mod1, mod2))



```

models 3 and 4 have worse FIC, rmse as expected 
FIC values here near those in book but not exactly
MLEs are the same, but book states
xbar - x  =  mean(ais$BMI) - 21.82 which is not 1.141 for females as stated in book 
presumably also algebraic vs numeric calculations somewhere, e.g. info matrix?


## Median at particular cov value, skewness may affect this. should pick biggest model. 

## Third central moment -- pick skewed model if degree of skewness sufficiently big 


todo should the focus function (mean) involve the error term for skew models?

```{r}

inds <- rbind(c(1,0,1,0), c(1,1,1,0), c(1,0,1,1), c(1,1,1,1))


```
